---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
* The data is unzipped and loaded using the read.csv. 
* Note: It is assumed that you have already downloaded data.

```{r}
unzip(zipfile="activity.zip")
Data <- read.csv('activity.csv',colClasses = c("integer", "Date", "integer"))
head(Data)
```
## What is mean total number of steps taken per day?

* The total number of steps taken per day is calculated as per following using tapply function.
```{r}
totalsteps <- tapply(Data$steps, Data$date, sum, na.rm=TRUE)
head(totalsteps)
```

* Histogram is of the total number of steps taken each day is as follow:
```{r warning=FALSE}
library(ggplot2)
qplot(totalsteps, xlab='Total steps each day', ylab = 'Frequency using binwith 1000', binwidth=1000)
```

* The mean and median of the total number of steps taken per day is as follow:

```{r}
stepsperDayMean <- mean(totalsteps)
stepsperDayMean

stepsperDayMedian <- median(totalsteps)
stepsperDayMedian
```

## What is the average daily activity pattern?
* A time series plot of the 5 minute interval and average number of steps taken, averaged across all days.
```{r}
intervalAvgData <- aggregate(x=list(meanSteps=Data$steps), by=list(interval=Data$interval), FUN=mean, na.rm=TRUE)

ggplot(data=intervalAvgData, aes(x=interval, y=meanSteps))+
  geom_line() + 
  xlab("5 minute interval")+
  ylab("avg number of steps taken")
```

* To determine which 5-minute interval on average across all days in the dataset contains the maximum number of steps will use which.max()

```{r}
maxnumofsteps<-intervalAvgData[which.max(intervalAvgData$meanSteps),]
maxnumofsteps
```

* The interval 835 has the maximum number of steps on average(206 steps).

## Imputing missing values

* The total number of missing values in the dataset will be given as:

```{r}
sum(is.na(Data$steps))
```

* Strategy to create a new data set is to replace a NA value with the mean.

```{r warning=FALSE, message=FALSE}
library(Hmisc)
ImputedData<-Data
ImputedData$steps<- impute(Data$steps, fun=mean)
```

* Histogram of the total number of steps taken each day with new data
```{r}
totalstepseachday <- tapply(ImputedData$steps, ImputedData$date, sum)
qplot(totalstepseachday, xlab='Total steps taken each day', ylab='Frequency using binwith=1000', binwidth=1000)
```

* Mean and Median of total number of steps taken per day.
```{r}
mean(totalstepseachday)

median(totalstepseachday)
```

* Without the missing data, the gap between mean and median is reduced is almost same.

## Are there differences in activity patterns between weekdays and weekends?

* A new factor variable in the dataset with two levels-"weekday" and "weekend" is created as follows:
```{r}
ImputedData$daytype <- ifelse(as.POSIXlt(ImputedData$date)$wday %in% c(0,6), 'weekend','weekday')

head(ImputedData)
```

* A time series plot of the full data is created with following R-functions.
```{r}
intervalAvgDataImputed<- aggregate(steps ~ interval + daytype, data=ImputedData, mean)

ggplot(intervalAvgDataImputed, aes(interval, steps))+
  geom_line()+
  facet_grid(daytype ~ .)+
  xlab("5 minute interval")+
  ylab("avg number of steps taken")
```