---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
* The data is unzipped and loaded using the read.csv. 
* Note: It is assumed that you have already downloaded data.

```{r}
unzip(zipfile="activity.zip")
Data <- read.csv('activity.csv',colClasses = c("integer", "Date", "integer"))
head(Data)
```
## What is mean total number of steps taken per day?

* The total number of steps taken per day is calculated as per following using tapply function.
```{r}
totalsteps <- tapply(Data$steps, Data$date, sum, na.rm=TRUE)
head(totalsteps)
```

* Histogram is of the total number of steps taken each day is as follow:
```{r warning=FALSE}
library(ggplot2)
qplot(totalsteps, xlab='Total steps each day', ylab = 'Frequency using binwith 1000', binwidth=1000)
```

* The mean and median of the total number of steps taken per day is as follow:

```{r}
stepsperDayMean <- mean(totalsteps)
stepsperDayMean

stepsperDayMedian <- median(totalsteps)
stepsperDayMedian
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
